NYC Spatial Models - Comparison of Spatial Models With and Without Twitter Data

### 2018-11-05 NYC_Twitter_SDM_models - Copy.Rmd
Reorganize model estimation and impacts comparison

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(group = "Long distance, no tweets", weight = 1, color = ~pal1(park_ls1)) %>%
  addPolygons(group = "Long distance, tweets", weight = 1, color = ~pal2(park_ls2)) %>%
  addPolygons(group = "Short distance, no tweets", weight = 1, color = ~pal4(park_ls4)) %>%
  addPolygons(group = "Short distance, tweets", weight = 1, color = ~pal5(park_ls5)) %>%
  addLayersControl(baseGroups =  c("Short distance, no tweets", 
                                   "Short distance, tweets", 
                                   "Long distance, no tweets", 
                                   "Long distance, tweets"))
```

A
Create regression formulas:
```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
#base_formula <- formula(~log(Pop_Density) + FulltimeWork + 
    #CollegeDeg + Single_Percent + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)
base_formula_age <- formula(~log(Pop_Density) + FulltimeWork + 
    CollegeDeg + Single_Percent + Pct0to17 + Pct18to29 + Pct30to64 + Pct65plus + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)

parks_formula_1a <- update(base_formula_age, . ~ . + park_ls1 )
parks_formula_2a <- update(base_formula_age, . ~ . + park_ls2 )
parks_formula_3a <- update(base_formula_age, . ~ . + park_ls4 )
parks_formula_4a <- update(base_formula_age, . ~ . + park_ls5 )

@@

Create a spatial weights matrix:

```{r W, echo=FALSE}
neighbors <- poly2nb(tracts, queen = FALSE)
W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
print(W, zero.policy = TRUE)
```

Estimate spatial models for the `OBESITY` variable. Model 1 of each type does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:

```{r spmodels obesity, echo=FALSE, message=FALSE, warning=FALSE}
obese_sar_age <- lagsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
obese_sar2_age <- lagsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data,
                      listw = W, zero.policy = TRUE)
obese_sar3_age <- lagsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data,
                     listw = W, zero.policy = TRUE)
obese_sem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
obese_sem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
obese_sem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
obese_sdm_age <- lagsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm2_age <- lagsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm3_age <- lagsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)

@@

Compare the models with `htmlreg()` in
the html output. First, the obesity models.

```{r obesitysummaryage, results="asis", echo=FALSE}
obesity_models_age <- list(obese_sar_age, obese_sar2_age, obese_sar3_age, obese_sem_age, obese_sem2_age, obese_sem3_age, obese_sdm_age, obese_sdm2_age, obese_sdm3_age, obese_sdem_age, obese_sdem2_age, obese_sdem3_age)
names(obesity_models_age) <-  paste(rep(c("SAR", "SEM" , "SDM", "SDEM"), each = 3),
                                rep(c("1", "2", "3")))
htmlreg(obesity_models_age, caption = "NYC: Spatial Models of Obesity Prevalence")
```

Estimate the impacts for `obese_sdm_age`.
```{r obese_sdm_age_impacts}

listw <- as(W, "CsparseMatrix")
trMC <- trW(listw, type="MC")
set.seed(1)
#summary(impacts(obese_sdm_age, tr=trMC, R=1000), zstats=TRUE)

>>>>

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(group = ls_names[1], weight = 1, color = ~pal1(park_ls1)) %>%
  addPolygons(group = ls_names[2], weight = 1, color = ~pal2(park_ls2)) %>%
  addPolygons(group = ls_names[3], weight = 1, color = ~pal4(park_ls4)) %>%
  addPolygons(group = ls_names[4], weight = 1, color = ~pal5(park_ls5)) %>%
  addLayersControl(baseGroups = ls_names)
```

# Obesity model estimation

Create a spatial weights matrix. It's showing that we have three regions with
no links, so we need to ensure that these are estimated in the model.

```{r W, echo=FALSE}
neighbors <- poly2nb(tracts, queen = TRUE)
W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
print(W, zero.policy = TRUE)
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
```

```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
base_formula <- formula( 
  ~ log(Pop_Density) + 
    FulltimeWork + CollegeDeg + Single_Percent + 
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity  
    PctBlack + PctAsian + PctOther + PctHispanic)

parks_formula_1 <- update(base_formula, . ~ . + park_ls1 )
parks_formula_2 <- update(base_formula, . ~ . + park_ls2 )
parks_formula_4 <- update(base_formula, . ~ . + park_ls4 )
parks_formula_5 <- update(base_formula, . ~ . + park_ls5 )
```

First, let's estimate the base model as an LM, SDM, and SEM model

```{r basemodels}
obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = tracts@data)
obese_base_sar <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sem <- errorsarlm(update(base_formula, OBESITY ~ .), 
                             data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE, 
                           type = "mixed")
```

A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
only going forward.

```{r lrtest}
lmtest::lrtest(obese_base_sem, obese_base_sdm)
```

```{r spmodels obesity, echo=FALSE, message=FALSE, warning=FALSE}
obese_ls1_sdm <- lagsarlm(update(parks_formula_1, OBESITY ~ .), 
                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls2_sdm <- lagsarlm(update(parks_formula_2, OBESITY ~ .), 
                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls4_sdm <- lagsarlm(update(parks_formula_4, OBESITY ~ .), 
                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls5_sdm <- lagsarlm(update(parks_formula_5, OBESITY ~ .), 
                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")

obesity_models <- list(obese_base_sdm, obese_ls1_sdm, obese_ls2_sdm, obese_ls4_sdm, obese_ls5_sdm)
names(obesity_models) <- c("Base", ls_names)
```

```{r}
screenreg(obesity_models)
```

Estimate the impacts for `obese_sdm_age`.

```{r obese_sdm_age_impacts}
impacts_extractor <- function(sdm) {

  impacts_summary <- summary(impacts(sdm, tr=trMC, R=1000), zstats = TRUE)

  coef <- list(
    direct = impacts_summary$direct_sum,
    indirect = impacts_summary$indirect_sum,
    total = impacts_summary$total_sum
  ) %>%
  lapply(function(s) {
    data_frame(
      var = names(s$statistics[,1]),
      impact = s$statistics[,1]
    )
  }) %>%
    bind_rows(.id = "effect")

  pval <- as_data_frame(impacts_summary$pzmat) %>%
    mutate(
     var = rownames(impacts_summary$pzmat)
    ) %>%
    gather("effect", "p-val", -var) %>%
    mutate(effect = tolower(effect))

  left_join(coef, pval, by = c("effect", "var"))
}

obesity_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(model, var, effect, output = paste(round(impact, 5), gtools::stars.pval(`p-val`))) %>%
  spread(model, output)

@@

@@
